---
title: "Multivariable Regression - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object into a dataframe. Fit a linear model of driver deaths with `kms` and `PetrolPrice` as predictors. Interpret your results.

```{r, Q1Ans}
library(datasets)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
fit <- lm(DriversKilled ~ kms + PetrolPrice, sb)
round(summary(fit)$coef, 4)
# Numbers need to be centered and maybe scaled
summary(sb$kms)
summary(sb$PetrolPrice)
library(dplyr)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms/1000,
             mmc=mm - mean(mm))
fit2 <- lm(DriversKilled ~ pp + mmc, sb)
round(summary(fit2)$coef, 4)
```

**ANSWER:** There are on average 123 drivers killed for the average gas price & distance driven. for every 1 standard deviation increase in gas price, we expect 7 to 8 fewer deaths when holding the distance traveled constant. For every 1000 km driven (holding gas prices constant). All results are significant w/ 95% confidence.

---

## Question 2

Predict the number of driver deaths at the average `kms` and `PetrolPrice` levels.

```{r, Q2Ans}
fit <- lm(DriversKilled ~ kms + PetrolPrice, sb)
predict(fit, newdata=data.frame(kms=mean(sb$kms), PetrolPrice=mean(sb$PetrolPrice)))
```

**ANSWER:** We predict 122 - 123 driver deaths at the average `kms` and `PetrolPrice`.

---

## Question 3

Take the residual for DriversKilled having regressed out `kms` and an intercept and the residual for `PetrolPrice` having regressed out `kms` and an intercept. Fit a regression through the origin of the two residuals and show that it is the same as your coefficient obtained in question 1

```{r, Q3Ans}
fitFull <- lm(DriversKilled ~ kms + PetrolPrice, sb)
sb$edk <- resid(lm(DriversKilled ~ kms, sb))
sb$epp <- resid(lm(PetrolPrice ~ kms, sb))
fitResid <- lm(edk ~ epp - 1, sb)
round(summary(fitFull)$coef, 4)
round(summary(fitResid)$coef, 4)
```

**ANSWER:** The coefficients are the same (-643.7895) with some (expected) rounding error.

---

## Question 4

Take the residual for `DriversKilled` having regressed out `PetrolPrice` and an intercept. Take the residual for `kms` having regressed out `PetrolPrice` and an intercept. Fit a regression through the origin of the two residuals and show that it is the same as your coefficient obtained in question 1.

```{r, Q4Ans}
sb$edk <- resid(lm(DriversKilled ~ PetrolPrice, sb))
sb$ekms <- resid(lm(kms ~ PetrolPrice, sb))
fitResid <- lm(edk ~ ekms - 1, sb)
round(summary(fitFull)$coef, 4)
round(summary(fitResid)$coef, 4)
```

**ANSWER:** The coefficients are the same (-0.0017) with some (expected) rounding error.