---
title: "Residuals and Residual Variation - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
# make this an external chunk that can be included in any file
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
knit_hooks$set(plot = knitr:::hook_plot_html)
```

## Question 1

Fit a linear regression model to the `father.son` dataset th the father as the predictor and the son as the outcome. Plot the father's height (horizontal axis) versus the residuals (vertical axis).

**ANSWER:** See plot.

```{r, Q1Ans}
library(UsingR)
library(ggplot2)
data(father.son)
fit <- lm(sheight ~ fheight, father.son)
father.son$resid <- fit$residuals
ggplot(father.son, aes(x=fheight, y=resid)) +
    geom_point(size=5, color="darkgreen", alpha=0.5) +
    geom_hline(yintercept=0, color="blue")
```

---

## Question 2

Refer to question 1. Directly estimate the residual variance and compare this estimate to the output of lm

**ANSWER:** See R code.

```{r, Q2Ans}
n <- nrow(father.son)
s1 <- summary(fit)$sigma
s2 <- sqrt(sum(father.son$resid^2) / (n - 2))
s1
s2
all.equal(s1, s2)
```

---

## Question 3

Refer to question 1. Give the $R^2$ for this model

```{r, Q3Ans}
summary(fit)$r.squared
```

**ANSWER:** 0.2513401

---

## Question 4

Load the `mtcars` dataset. Fit a linear regression with miles per gallon as the outcome and horsepower as the predictor. Plot horsepower versus the residuals.

**ANSWER:** See plot. The curvature suggests linear may not be the best fit.

```{r, Q4Ans}
data(mtcars)
fit <- lm(mpg ~ hp, mtcars)
mtcars$resid <- fit$residuals
ggplot(mtcars, aes(x=hp, y=resid)) +
    geom_point(size=5, color="darkgreen", alpha=0.5) +
    geom_hline(yintercept=0, "blue")
```

---

## Question 5

Refer to question 4. Directly estimate the residual variance and compare this estimate to the output of lm.

**ANSWER:** See R code.

```{r, Q5Ans}
n <- nrow(mtcars)
s1 <- summary(fit)$sigma
s2 <- sqrt(sum(mtcars$resid^2) / (n - 2))
s1
s2
all.equal(s1, s2)
```

---

## Question 6

Refer to question 4. Give the $R^2$ for this model

```{r, Q6Ans}
summary(fit)$r.squared
```

**ANSWER:** 0.6024373