---
title: "Inference in Regression - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Test whether the slope coefficient for the father.son data is different from zero (father as predictor, son as outcome)

$$
H_0 : \hat\beta_1 = 0 \\
H_\alpha : \hat\beta_1 \ne 0
$$

```{r, Q1Ans}
library(UsingR)
data(father.son)
y <- father.son$sheight
x <- father.son$fheight
fs.fit <- lm(y ~ x)
fs.coef <- summary(fs.fit)$coefficients
fs.coef
```

**ANSWER:** $\hat\beta_1$ is 0.514, with a $p$-value of 0. Therefore we reject the null hypothesis

---

## Question 2

Refer to question 1. Form a confidence interval for the slope coefficient.

```{r, Q2Ans}
confint(fs.fit)[2,]
```

**ANSWER:** CI at 95% confidence is 0.461 to 0.567

---

## Question 3

Refer to question 1. Form a confidence interval for the intercept (center the father's height first to get an intercept that is easier to interpret)

```{r, Q3Ans}
fs.fit.c <- lm(y ~ I(x - mean(x)))
confint(fs.fit.c)[1,]
```

**ANSWER:** The son's height around the mean father's height has a CI at 95% confidence of 68.538 to 68.830

---

## Question 4

Refer to question 1. Form a mean value interval for the expected son's height at the average father's height.

```{r, Q4Ans}
predict(fs.fit, newdata=data.frame(x=mean(x)), interval="confidence")
```

**ANSWER:** 68.684 w/ CI at 95% of 68.538 to 68.830

---

## Question 5

Refer to question 1. Form a prediction interval for the son's height at the average father's height.

```{r, Q5Ans}
predict(fs.fit, newdata=data.frame(x=mean(x)), interval="prediction")
```

**ANSWER:** 68.684 w/ CI at 95% of 63.901 and 73.467

---

## Question 6

Load the `mtcars` dataset. Fit a linear regression with miles per gallong as the outcome and horsepower as the predictor. Test whether or not the horsepower coefficient is statistically different from zero. Interpret your test.

$$
H_0 : \hat\beta_1 = 0 \\
H_\alpha : \hat\beta_1 \ne 0
$$

```{r, Q6Ans}
data(mtcars)
y <- mtcars$mpg
x <- mtcars$hp
cars.fit <- lm(y ~ x)
summary(cars.fit)$coefficients
```

**ANSWER:** $\hat\beta_1$ is -0.068, with a $p$-value of 0. Therefore we reject the null hypothesis.

---

## Question 7

Refer to question 6. Form a confidence interval for the slope coefficient.

```{r, Q7Ans}
confint(cars.fit)[2,]
```

**ANSWER:** Slope CI at 95% is -0.089 to -0.048

---

## Question 8

Refer to question 6. Form a confidence interval for the intercept (center the HP variable first)

```{r, Q8Ans}
cars.fit.c <- lm(y ~ I(x - mean(x)))
confint(cars.fit.c)[1,]
```

**ANSWER:** (Centered) Intercept CI at 95% is 18.696 to 21.485.

---

## Question 9

Refer to question 6. Form a mean value interval for the expected MPG for the average HP.

```{r, Q9Ans}
predict(cars.fit, newdata=data.frame(x=mean(x)), interval="confidence")
```

**ANSWER:** 20.091 w/ CI at 95% of 18.696 to 21.485

---

## Question 10

Refer to question 6. Form a prediction interval for the expected MPG for the average HP.

```{r, Q10Ans}
predict(cars.fit, newdata=data.frame(x=mean(x)), interval="prediction")
```

**ANSWER:** 20.091 w/ CI at 95% of 12.079 to 28.102.

---

## Question 11

Refer to question 6. Create a plot that has the fitted regression line plus curves at the expected values and prediction interval.

**ANSWER:** See plot

```{r, Q11Ans, echo=FALSE, fig.height=5, fig.width=5}
library(ggplot2)
newx <- data.frame(x=seq(min(x), max(x), length=100))
p1 <- data.frame(predict(cars.fit, newdata=newx, interval="confidence"))
p1$interval <- "Confidence"
p1$x <- newx$x
p2 <- data.frame(predict(cars.fit, newdata=newx, interval="prediction"))
p2$interval <- "Prediction"
p2$x <- newx$x
dat <- rbind(p1, p2)
ggplot(dat, aes(x=x, y=fit)) +
    geom_ribbon(aes(ymin=lwr, ymax=upr, fill=interval), alpha=0.2) +
    geom_line()
```