---
title: "Statistical Linear Regression Models - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
# make this an external chunk that can be included in any file
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
knit_hooks$set(plot = knitr:::hook_plot_html)
```

## Question 1

Fit a linear regression model to the `father.son` dataset with the father as the predictor and the son as the outcome. Give a $p$-value for the slope coefficient and perform the relevant hypothesis test.

$$
ht_{son} = \beta_0 + \beta_1 ht_{father} + \epsilon_i \\
H_0 : \beta_1 = 0 \\
H_\alpha : \beta_1 \ne 0
$$

```{r, Q1Ans}
library(UsingR)
data(father.son)
fit <- lm(sheight ~ fheight, data=father.son)
summary(fit)$coef
```

**ANSWER:** Our fit estimates $\beta_1$ as 0.514 with a very high $t$-value (19.006) and a very low $p$-value (almost 0). Therefore the null hypothesis is rejected.

---

## Question 2

Refer to question 1. Interpret both parameters. Recenter for the intercept if necessary.

```{r, Q2Ans}
fit.c <- lm(sheight ~ I(fheight - mean(fheight)), data=father.son)
summary(fit.c)$coef
```

**ANSWER:** The son's height for the average height of all fathers is 68.684. For every inch of father's height, we can predict 0.514 inches of the son's height.

---

## Question 3

Refer to question 1. Predict the son's height if the father's height is 80 niches. Would you recommend this prediction? Why or why not?

```{r, Q3Ans}
predict(fit, newdata=data.frame(fheight=80))
summary(father.son)
```

**ANSWER:** 75. This is not a reliable prediction, as the maximum heights in our sample set is lower than the value we're trying to predict. We have no estimate of variability at that height.

---

## Question 4

Load the `mtcars` dataset. Fit a linear regression with miles per gallon as the outcome and horsepower as the predictor. Interpret your coefficients, recenter for the intercept if necessary.

```{r, Q4Ans}
data(mtcars)
summary(lm(mpg ~ I(hp - mean(hp)), data=mtcars))$coef
```

**ANSWER:** The `mpg` for a car w/ average `hp` is 20.091. For every unit of gross horsepower, we predict a -0.068 decrease in miles per gallon. This result is highly significant, with a $p$-value of nearly 0.

---

## Question 5

Refer to question 4. Overlay the fit onto a scatterplot.

**ANSWER:** See plot.

```{r, Q5Ans}
library(ggplot2)
ggplot(mtcars, aes(x=hp, y=mpg)) +
    geom_point(cex=5, color="darkgreen", alpha=0.5) +
    geom_smooth(method="lm", se=F, lwd=2)
```

---

## Question 6

Refer to question 4. Test they hypothesis of no linear relationship between horsepower and miles per gallon.

$$
mpg = \beta_0 + \beta_1 hp + \epsilon_i \\
H_0 : \beta_1 = 0 \\
H_\alpha : \beta_1 \ne 0
$$

```{r, Q6Ans}
summary(lm(mpg ~ hp, data=mtcars))$coef
```

**ANSWER:** With a $\beta_1$ of -0.068 and a $p$-value of nearly 0, the null hypothesis of no linear relationship is rejected.

---

## Question 7

Refer to question 4. Predict the miles per gallong for a horsepower of 111

```{r, Q7Ans}
summary(mtcars[,c("mpg","hp")])
fit <- lm(mpg ~ hp, data=mtcars)
predict(fit, newdata=data.frame(hp=111))
```

**ANSWER:** 22.526. Both the predicted mpg & the hp of 111 are reasonable values given our data.