---
title: "Residuals, Diagnostics, & Variation - Exercises"
output: 
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Fit a linear model of driver deaths w/ `kms`, `PetrolPrice`, and `law` as predictors

**ANSWER:** See R code below.

```{r, Q1Ans}
library(datasets)
library(dplyr)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms / 1000,
             mmc=mm - mean(mm))
fit <- lm(DriversKilled ~ mmc + pp + law, sb)
summary(fit)
```

---

## Question 2

Refer to question 1. Directly estimate the residual variation via the function `resid`. Compare with R's residual variation estimate.

**ANSWER:** See R code below.

```{r, Q2Ans}
summary(fit)$sigma^2
sum(resid(fit)^2)/fit$df
```

---

## Question 3

Refer to question 1. Perform an analysis of diagnostic measures including dffits, dfbetas, influence, and hat diagonals.

```{r, Q3Ans}
plot(fit)
plot(dffits(fit))
plot(dfbetas(fit)[,2])
plot(dfbetas(fit)[,3])
plot(dfbetas(fit)[,4])
plot(cooks.distance(fit))
plot(hatvalues(fit))
```

**ANSWER:** The observations toward the end of the frame deserve some greater scrutiny, based on the hat values & cooks distance for those records. In particular, `law` may be involved, based on the influence those observations have on $\beta_{law}$. The first observation's `PetrolPrice` also could be scrutinized.