---
title: "Multivariable Regression Examples - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Do exercise 1 of the previous chapter if you haven't already. Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Fit a linear model of driver deaths with `kms` and `PetrolPrice`as predictors. Interpret your results.

```{r, Q1Ans}
library(datasets)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
fit <- lm(DriversKilled ~ kms + PetrolPrice, sb)
round(summary(fit)$coef, 4)
# Numbers need to be centered and maybe scaled
summary(sb$kms)
summary(sb$PetrolPrice)
library(dplyr)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms/1000,
             mmc=mm - mean(mm))
fit2 <- lm(DriversKilled ~ pp + mmc, sb)
round(summary(fit2)$coef, 4)
```

**ANSWER:** There are on average 123 drivers killed for the average gas price & distance driven. For every 1 standard deviation increase in gas price, we expect 7 to 8 fewer deaths when holding the distance traveled constant. For every 1000 km driven (holding gas prices constant) we expect 1 to 2 fewer deaths. All results are significant w/ 95% confidence.

---

## Question 2

Repeat question 1 for the outcome being the log of the count of driver deaths. Interpret your coefficients.

$$
\begin{eqnarray*}
log\left(Y_i\right) & = & \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \epsilon_i \\
E\left[log\left(Y_i\right)\right] & = & \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} \\
e^{E\left[log\left(Y_i\right)\right]} & = & e^{\beta_0} + e^{\beta_1 X_{1i}} + e^{\beta_2 X_{2i}}
\end{eqnarray*}
$$

```{r, Q2Ans}
fit3 <- lm(I(log(DriversKilled)) ~ pp + mmc, sb)
1 - exp(summary(fit3)$coef[2,1])
1 - exp(summary(fit3)$coef[3,1])
```

**ANSWER:** For every 1 standard deviation increase in gas price, we expect a 6% decrease in driver deaths when holding the distance traveled constant. For every 1000 km driven (holding gas prices constant) we expect a 1% decrease in deaths. All results are significant w/ 95% confidence.

---

## Question 3

Refer to question 1. Add the dummy variable `law` and interpret the results. Repeat this question with a factor variable that you create called `lawFactor` that takes the levels `No` and `Yes`. Change the reference level from `No` to `Yes`.

```{r, Q3Ans}
fit4 <- lm(DriversKilled ~ pp + mmc + law, sb)
round(summary(fit4)$coef, 4)
sb$lawFactor <- relevel(factor(sb$law, labels=c("No","Yes")), "Yes")
fit5 <- lm(DriversKilled ~ pp + mmc + lawFactor, sb)
round(summary(fit5)$coef, 4)
```

**ANSWER:** The law being in effect resulted in 11 to 12 fewer driver deaths on average, w/ 95% confidence, holding gas prices & distance driver as constant.

---

## Question 4

Discretize the `PetrolPrice` variable into four factor levels. Fit the linear model with this factor to see how R treats multiple level factor variables.

**ANSWER:** See R code below.

```{r, Q4Ans}
sb$ppf <- as.factor((sb$pp <= -1.5) + (sb$pp <= 0) +
                        (sb$pp <= 1.5) + (sb$pp < Inf))
fit6 <- lm(DriversKilled ~ ppf + mmc + law, sb)
round(summary(fit6)$coef, 4)
```

---

## Question 5

Perform the plot requested at the end of the last chapter.

```{r, Q5Ans}
require(ggplot2)
require(dplyr)
data(swiss)
swiss <- mutate(swiss, CatholicBin = 1 * (Catholic > 50))
fit <- lm(Fertility ~ Agriculture * factor(CatholicBin), swiss)
ggplot(swiss, aes(x=Agriculture, y=Fertility, color=factor(CatholicBin))) +
    geom_point(size=6, color="black") +
    geom_point(size=4) +
    xlab("% in Agriculture") +
    ylab("Fertility") +
    geom_abline(intercept=coef(fit)[1], slope=coef(fit)[2], size=2, 
                color="salmon") +
    geom_abline(intercept=coef(fit)[1] + coef(fit)[3],
                slope=coef(fit)[2] + coef(fit)[4], size=2,
                color=hcl(h=195, c=100, l=65))
```