---
title: "Multivariable Regression - Adjustment - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Fit a linear model of driver deaths with `kms` and `PetrolPrice` as predictors. Interpret your results.

```{r, Q1Ans}
library(datasets)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
fit <- lm(DriversKilled ~ kms + PetrolPrice, sb)
round(summary(fit)$coef, 4)
# Numbers need to be centered and maybe scaled
summary(sb$kms)
summary(sb$PetrolPrice)
library(dplyr)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms/1000,
             mmc=mm - mean(mm))
fit2 <- lm(DriversKilled ~ pp + mmc, sb)
round(summary(fit2)$coef, 4)
```

**ANSWER:** There are on average 123 drivers killed for the average gas price & distance driven. For every 1 standard deviation increase in gas price, we expect 7 to 8 fewer deaths when holding the distance traveled constant. For every 1000 km driven (holding gas prices constant) we expect 1 to 2 fewer deaths. All results are significant w/ 95% confidence.

---

## Question 2

Compare the `kms` coefficient with and without the inclusion of the `PetrolPrice` variable in the model.

```{r, Q2Ans}
cor(sb$pp, sb$mmc)
fit3 <- lm(DriversKilled ~ mmc, sb)
summary(fit2)$coef
summary(fit3)$coef
anova(fit2, fit3)
```

**ANSWER:** Holding gas price constant, distance driven is related to a decrease of 1 to 2 driver deaths per 1000 km driven, with a $p$-value of .0049. When not fixing gas prices, distance driven is related to a decrease of 2 to 3 drivers deaths, with a $p$-value of effectively 0. Correlation between distance driven & gas prices suggest some interaction here.

---

## Question 3

Compare the `PetrolPrice` coefficient with and without the inclusion of the `kms` variable in the model.

```{r, Q3Ans}
cor(sb$pp, sb$mmc)
fit4 <- lm(DriversKilled ~ pp, sb)
summary(fit2)$coef
summary(fit4)$coef
anova(fit2, fit4)
```

**ANSWER:** Holding distance driven as constant, gas prices are related to a decrease of 7 to 8 driver deaths per 1 std deviation change in gas prices, with a $p$-value of .00002. When not fixing distance driven, gas prices are related to a decrease of 9 to 10 drivers deaths, with a $p$-value of effectively 0. Correlation between distance driven & gas prices suggest some interaction here.

---

## Just for fun

```{r}
fit5 <- lm(DriversKilled ~ pp + mmc + pp * mmc, sb)
summary(fit5)$coef
anova(fit2, fit3, fit4, fit5)
```