---
title: "Multivariable Regression - Adjustment - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Fit a linear model of driver deaths with `kms` and `PetrolPrice` as predictors. Interpret your results.

```{r, Q1Ans}
library(datasets)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
fit <- lm(DriversKilled ~ kms + PetrolPrice, sb)
round(summary(fit)$coef, 4)
# Numbers need to be centered and maybe scaled
summary(sb$kms)
summary(sb$PetrolPrice)
library(dplyr)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms/1000,
             mmc=mm - mean(mm))
fit2 <- lm(DriversKilled ~ pp + mmc, sb)
round(summary(fit2)$coef, 4)
```

**ANSWER:** There are on average 123 drivers killed for the average gas price & distance driven. For every 1 standard deviation increase in gas price, we expect 7 to 8 fewer deaths when holding the distance traveled constant. For every 1000 km driven (holding gas prices constant) we expect 1 to 2 fewer deaths. All results are significant w/ 95% confidence.

---

