---
title: "Multiple Variables and Model Fit - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Fit a linear model of driver deaths with `kms`, `PetrolPrice`, and `law` as predictors.

```{r, Q1Ans}
require(datasets)
require(dplyr)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms / 1000,
             mmc=mm - mean(mm))
fit <- lm(DriversKilled ~ pp + mmc + law, sb)
summary(fit)
```

---

## Question 2

Perform a model selection exercise to arrive at a final model.

```{r, Q2Ans}
library(car)
vif(fit)
fit0 <- update(fit, DriversKilled ~ law)
fit1 <- update(fit, DriversKilled ~ law + mmc)
fit2 <- update(fit, DriversKilled ~ law + pp)
fit3 <- update(fit, DriversKilled ~ law + mmc + pp)
anova(fit0, fit1, fit2, fit3)
rbind(summary(fit0)$coef[2,],
      summary(fit1)$coef[2,],
      summary(fit2)$coef[2,],
      summary(fit3)$coef[2,])
anova(fit0, fit3)
shapiro.test(fit0$residuals)
shapiro.test(fit3$residuals)
```

**ANSWER:** A model which includes `law`, `kms`, and `PetrolPrice` is the more effective model given the available data. However, given that our residuals fail tests for normality at 95%, there are concerns about the model fit. 