---
title: "Poisson Regression - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Fit a Poisson regression GLM with `DriversKilled` as the outcome and `kms`, `PetrolPrice`, and `law` as predictors. Interpret your results

```{r, Q1Ans}
require(datasets)
require(dplyr)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
sb <- mutate(sb,
             pp=(PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm=kms / 1000,
             mmc=mm - mean(mm))
fit <- glm(DriversKilled ~ pp + mmc + law, poisson, sb)
summary(fit)
exp(confint(fit))
anova(fit)
```

---

## Question 2

Refer to question 1. Fit a linear model with the log of drivers killed as the outcome. Interpret your results.

```{r, Q2Ans}
fit2 <- lm(log(DriversKilled) ~ pp + mmc + law, sb)
round(summary(fit)$coef, 3)
round(summary(fit2)$coef, 3)
```

---

## Question 3

Refer to question 1. Fit your Poisson log-linear model with `drivers` as a log offset (to consider the proportion of drivers killed of those killed or seriously injured)

```{r, Q3Ans}
fit3 <- glm(DriversKilled ~ pp + mmc + law, poisson, sb, offset=log(drivers))
summary(fit3)
```

---

## Question 4

Refer to Question 1. Use the `anova` function to compare the models with just `law`, `law` and `PetrolPrice`, and all three predictors.

```{r, Q4Ans}
anova(glm(DriversKilled ~ law + pp + mmc, poisson, sb), "chisq")
```