---
title: "Logistic Regression - Exercises"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Question 1

Load the dataset `Seatbelts` as part of the `datasets` package via `data(Seatbelts)`. Use `as.data.frame` to convert the object to a dataframe. Create a new outcome variable for whether or not greater than 119 drivers were killed that month. Fit a logistic regression GLM with this variable as the outcome and `kms`, `PetrolPrice`, and `law` as predictors. Interpret your parameters.

```{r, Q1Ans}
require(datasets)
require(dplyr)
data(Seatbelts)
sb <- as.data.frame(Seatbelts)
sb <- mutate(sb,
             pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
             mm = kms / 1000,
             mmc = mm - mean(mm),
             dk119 = 1 * (DriversKilled > 119))
fit <- glm(dk119 ~ pp + mmc + law, binomial, sb)
1 - exp(summary(fit)$coef[,1])
exp(confint(fit))
anova(fit)
```

**ANSWER:** We saw a 45% decrease in the odds that > 119 drivers were killed once the law was enacted, holding gas prices & distance traveled constant.

---

## Question 2

Fit a binomial model with `DriversKilled` as the outcome and `drivers` as the total count with `kms`, `PetrolPrice`, and `law` as predictors, interpret your results.

```{r, Q2Ans}

```

---